\bfseries\LARGE\singlespacing{Transmission of airborne respiratory infections with and without air cleaners in a Swiss school under non-pandemic conditions in 2023: A modeling study of epidemiological, environmental, and molecular data}

### 2023-07-05 doc/paper/manuscript.tex
Updates from Overleaf

Our study indicated that air cleaners can reduce the risk of infection, although we we only were able tho show a trend. Other infection control measures, such as universal mask use, appear to be more effective (REFS). There may be several reasons for this. Compared to \eg mask wearing, air cleaners cannot influence transmission outside the classrooms. Furthermore, air cleaners are unable to prevent close range high particle density aerosol transmission. By contrast, face masks can also reduce the generation of airborne infectious particles and can thus mitigate transmission over short ranges. Evidence from our previous\cite{Banholzer2023PLoSMed} and this study further suggest that air cleaners are more effective in the removal of the larger particles ($>5\mu$m) than the smaller particles ($\leq5\mu$m). Yet, a larger proportion of respiratory viruses are carried in the smaller particles\cite{Fennelly2020}.

@@

We detected few respiratory viruses in bioaerosol samples (3~positive bioaoerosl and 50~positive saliva samples), especially when comparing to our previous study\cite{Banholzer2023PLoSMed} (10~positive bioaerosol and 21~positive saliva samples). The number of viruses detected on the filters of the air cleaners was also lower (this study: 4~positive samples in 160~swabs, previous study: 8~positive samples in 80~swabs). The bioaerosol sampling devices were the same in both studies and technical issues were not observed. It is also unlikely that environmental factors are responsible for the differences. Temperature and humidity were similar, and CO$_2$ levels were even higher in this study (1,702$\pm$370\,ppm in this vs 1,064$\pm$232\,ppm in the previous study), suggesting that classrooms were not as well ventilated as during the COVID-19 pandemic. Therefore, we think that the differences in the number of positive bioaerosols samples between both studies is most likely related to differences in the distribution of respiratory viruses. Strikingly, we detected a wide range of respiratory viruses (Flu B, HRV, AdV, etc) in this study, as compared to only SARS-CoV-2 in our previous study. The shift in the pattern of respiratory viruses has been observed in other studies (xxx). SARS-CoV-2 may be present in higher concentrations in air and may be more easily detected in bioaerosol samples due to differences in physicochemical properties of virus-laden aerosols (\eg physical size, viral load, and infectivity) that influence the generation and survival of airborne viral RNA\cite{Wang2021,Sattar2016Book}. In addition, although respiratory viruses have been detected in bioaerosols (\eg Flu\cite{Bischoff2013JID,Pan2017mSphere}, HRV\cite{Myatt2004AJRCCM,Fabian2011JAMPDD}, AdV\cite{Nguyen2017OFID,Pan2017mSphere}, and CoV\cite{Guo2020EID,Liu2020Nature}), bioaerosol sampling remains a technical challenge\cite{Wang2021 ,Belser2023PLOSPath,Bekking2019IORV,Mainelis2020AST}.

>>>>

Our study indicated that air cleaners can reduce the risk of infection, although we we only were able tho show a trend. Other infection control measures, such as universal mask use, appear to be more effective\cite{Banholzer2023PLoSMed,Heinsohn2022,Gettings2021}. There may be several reasons for this. Compared to \eg mask wearing, air cleaners cannot influence transmission outside the classrooms. Furthermore, air cleaners are unable to prevent close range high particle density aerosol transmission. By contrast, face masks can also reduce the generation of airborne infectious particles and can thus mitigate transmission over short ranges. Evidence from our previous\cite{Banholzer2023PLoSMed} and this study further suggest that air cleaners are more effective in the removal of the larger particles ($>5\mu$m) than the smaller particles ($\leq5\mu$m). Yet, a larger proportion of respiratory viruses are carried in the smaller particles\cite{Fennelly2020}.

@@

We detected few respiratory viruses in bioaerosol samples (3~positive bioaoerosl and 50~positive saliva samples), especially when comparing to our previous study\cite{Banholzer2023PLoSMed} (10~positive bioaerosol and 21~positive saliva samples). The number of viruses detected on the filters of the air cleaners was also lower (this study: 4~positive samples in 160~swabs, previous study: 8~positive samples in 80~swabs). The bioaerosol sampling devices were the same in both studies and technical issues were not observed. It is also unlikely that environmental factors are responsible for the differences. Temperature and humidity were similar, and CO$_2$ levels were even higher in this study (1,702$\pm$370\,ppm in this vs 1,064$\pm$232\,ppm in the previous study), suggesting that classrooms were not as well ventilated as during the COVID-19 pandemic. Therefore, we think that the differences in the number of positive bioaerosols samples between both studies is most likely related to differences in the distribution of respiratory viruses. Strikingly, we detected a wide range of respiratory viruses (Flu B, HRV, AdV, etc) in this study, as compared to only SARS-CoV-2 in our previous study. The shift in the pattern of respiratory viruses has been observed in other studies\cite{Nygaard2023Lancet,Sauteur2022EuroSurv}. SARS-CoV-2 may be present in higher concentrations in air and may be more easily detected in bioaerosol samples due to differences in physicochemical properties of virus-laden aerosols (\eg physical size, viral load, and infectivity) that influence the generation and survival of airborne viral RNA\cite{Wang2021,Sattar2016Book}. In addition, although respiratory viruses have been detected in bioaerosols (\eg Flu\cite{Bischoff2013JID,Pan2017mSphere}, HRV\cite{Myatt2004AJRCCM,Fabian2011JAMPDD}, AdV\cite{Nguyen2017OFID,Pan2017mSphere}, and CoV\cite{Guo2020EID,Liu2020Nature}), bioaerosol sampling remains a technical challenge\cite{Wang2021 ,Belser2023PLOSPath,Bekking2019IORV,Mainelis2020AST}.
